---
title: "Portfolio Analysis"
author: "Ashish Mathew"
format: html
---

```{r}
#| include: false
library(rvest)
library(dplyr)
library(quantmod)
library(ggplot2)
library(tidyquant)
library(tidyr)
library(ggcorrplot)
library(quantreg)
library(PerformanceAnalytics)
library(quadprog)
```

## Market Analysis
### Get Price Data
```{r}
# scrape list of stocks currently in the S&P 500
sp500_url = "https://en.wikipedia.org/wiki/List_of_S%26P_500_companies"
sp500_url = read_html(sp500_url)
data = sp500_url |> 
  html_element("table") |> 
  html_table()
colnames(data) = c('symbol','company','gics_sector','gics_sub_industry','headquarters','date_added','cik','founded')

# create a list of tickers
tickers = sort(data$symbol)
data |> head()
#data |> write.csv(file = "./data/sp500_list.csv", row.names = FALSE)
```

### Composition of the S&P500
```{r}
data |> 
  count(gics_sector) |> 
  ggplot(aes(x=reorder(gics_sector,n),y=n))+
  geom_bar(stat='identity')+
  coord_flip()+
  labs(title = "S&P500 Stocks by Sector",y="Count",x="Sector")
```

The S&P500 which tracks the performance of the 500 largest companies in
the US is weighted towards industrial, financial and technology stocks.

```{r}
price_data = tq_get(
  x=tickers,get='stock.prices',
  from='2023-01-01',to='2024-05-31') |> 
  left_join(y = data, by = 'symbol') |> 
  select(symbol, company, gics_sector, date_added, 
         founded, date, adjusted, volume)

price_data = price_data |> 
  group_by(symbol) |> 
  mutate(
    daily_return = (adjusted / lag(adjusted)) - 1,
    years_public = 2025 - as.integer(substr(date_added,1,4)),
    company_age = 2025 - as.integer(substr(founded,1,4))
  ) |> 
  filter(!is.na(daily_return)) |> 
  select(symbol, company, gics_sector, years_public, company_age,
         date, adjusted, volume, daily_return)
```

### Sector Correlation Matrix
```{r}
price_data |> 
  group_by(gics_sector, date) |>
  summarise(avg_return = mean(daily_return)) |> 
  pivot_wider(
    id_cols = "date", 
    names_from = "gics_sector", 
    values_from = "avg_return") |> 
  select(where(is.numeric)) |> 
  cor(use = "complete.obs") |> 
  ggcorrplot(
    method = "square", 
    type = "lower", 
    lab = 1, 
    lab_size = 2,
    colors = c("red","white","green"),
    title = "Sector Correlation Matrix")
```

From the correlation matrix, we can see that sectors like IT show a low
correlation with sectors like Energy, Consumer Staples and Utilities.
Historically this holds true since IT stocks are mainly growth stocks
that benefit when interest rates are low meanwhile Energy, Consumer
Staples and Utilities are more mature value stocks that have consistent
performance regardless of the interest-rate environment.

The risk of a specific stock can be influenced by several factors ranging from company specific risks to risks present in the country it operates in. Investors expect to be compensated for taking on additional risk with higher returns. 

Modern Portfolio Theory, seeks to minimize the risk of a portfolio through diversification, i.e. by picking a basket of uncorrelated assets to minimize variance and maximize risk-adjusted return.

$TotalRisk = Market Risk + IndustryRisk + Company Risk + CountryRisk$

-   Industry and Company risk can be minimized by holding a broad range
    of equities across various industries.
-   Market risk can be minimized by holding non-equity instruments like
    treasury bills and/or commodities like gold that act as a safe haven when markets are in decline.
-   Country risk is a bit more complicated, some advocate for owning
    companies that operate in different countries while critics point
    out that large US firms conduct business and have supply chains
    across geographies, so they should have low country specific risk.

For simplicity, this analysis only looks at companies listed in the US market.

```{r}
# Looking at top 4 stocks (based on Sharpe Ratio) in each GICS sector
price_data |> 
  group_by(symbol, gics_sector) |> 
  summarise(
    avg_daily_return = mean(daily_return),
    sd_daily_return = sd(daily_return)
  ) |> 
  mutate(
    sharpe_ratio = avg_daily_return / sd_daily_return # Assume Risk-Free Rate of 0%
  ) |> 
  group_by(gics_sector) |> 
  arrange(desc(sharpe_ratio), .by_group = T) |> 
  slice_head(n = 4)
```

We can create a portfolio by selecting top-performers across a range of
sectors. For this analysis we decided to construct a portfolio with the
below equities and indices.

**Equity** 
IT: DELL, NVDA, MU (Micron) 

Industrial: GE, UBER 

Financials: AXP, JPM 

Consumer Discretionary: AMZN, HLT 

Energy: TRGP 

Healthcare: LLY,BSX 

Utilities: NRG 

**Indices** 
Fixed Income: BND (Bond ETF) 

Commodities: GSG (Commodity ETF)

## Portfolio Analysis
### Get Price Data
```{r}
tickers = c(
    "DELL",
    "NVDA",
    "MU", # Tech
    "GE",
    "UBER", # Industrials
    "AXP",
    "JPM", # Financials
    "AMZN",
    "HLT", # Consumer Discretionary
    "TRGP", # Energy
    "LLY",
    "BSX", # Health Care
    "NRG", # Utilities
    "BND",
    "GSG" # Fixed-Income and Commodity ETFs
)
getSymbols(tickers, from = "2023-01-01", to = "2024-05-31")
prices = do.call(merge, lapply(tickers, function(t) Ad(get(t))))
returns = na.omit(Return.calculate(prices))

tickers = c("SPY") # S&P500 as benchmark
getSymbols(tickers, from = "2023-01-01", to = "2024-05-31")
spy_prices = do.call(merge, lapply(tickers, function(t) Ad(get(t))))
spy_returns = na.omit(Return.calculate(spy_prices))
```

### Equal Weight Portfolio
```{r}
n_assets = ncol(returns)
weights = rep(1 / n_assets, n_assets)
equal_weights = weights

portfolio_returns = Return.portfolio(returns, weights = weights)
common_dates = index(portfolio_returns)[index(portfolio_returns) %in% index(spy_returns)]
portfolio_returns_aligned = portfolio_returns[common_dates]
spy_returns_aligned = spy_returns[common_dates]
eq_weight_returns_aligned = portfolio_returns_aligned

combined_returns= merge.xts(portfolio_returns_aligned, spy_returns_aligned)
colnames(combined_returns) = c("Eq. Weight Portfolio", "S&P 500")

chart.CumReturns(combined_returns,
                 wealth.index = T,
                 main = "Portfolio vs S&P500",
                 legend.loc = "topleft")

chart.Drawdown(combined_returns,
               legend.loc = "bottomleft",
               main = "Drawdowns")

SharpeRatio.annualized(combined_returns, Rf = 0)
```
An equal weighted portfolio across sectors offers nearly twice the market return for similar draw down leading to a 2x difference in Sharpe Ratio (risk-adjusted return)

### Minimum Variance Portfolio
```{r}
cov_matrix = cov(returns)
n_assets = ncol(returns)

Dmat = 2 * cov_matrix
dvec = rep(0, n_assets)

# Constraints: sum of weights = 1, weights >= 0 (no short selling)
Amat = cbind(rep(1, n_assets), diag(n_assets))
bvec = c(1, rep(0, n_assets))

# Solve QP
sol = solve.QP(Dmat, dvec, Amat, bvec, meq = 1)

# Optimal weights
opt_weights = sol$solution
names(opt_weights) = colnames(returns)
min_var_weights = opt_weights
portfolio_returns = Return.portfolio(returns, weights = opt_weights)

common_dates = index(portfolio_returns)[index(portfolio_returns) %in% index(spy_returns)]

portfolio_returns_aligned = portfolio_returns[common_dates]
spy_returns_aligned = spy_returns[common_dates]
min_var_returns_aligned = portfolio_returns_aligned

combined_returns= merge.xts(portfolio_returns_aligned, spy_returns_aligned)
colnames(combined_returns) = c("Portfolio", "S&P 500")

chart.CumReturns(combined_returns,
                 wealth.index = T,
                 main = "Portfolio vs S&P500",
                 legend.loc = "topleft")

chart.Drawdown(combined_returns,
               legend.loc = "bottomleft",
               main = "Drawdowns")

SharpeRatio.annualized(combined_returns, Rf = 0)
```

While the optimized portfolio has a lower annualized return, it also shows limited draw down, offering a less volatile growth rate compared to the broader market

### Risk Adjusted Return Maximized Portfolio
```{r}
mean_returns = colMeans(returns)
n_assets = ncol(returns)

# Dummy Dmat and dvec for linear return optimization
Dmat_dummy = diag(n_assets)
dvec_dummy = rep(0, n_assets)

Amat = cbind(
  rep(1, n_assets),     # sum of weights = 1
  diag(n_assets)        # weights >= 0
)
bvec = c(1, rep(0, n_assets))

# Min return
result_min = solve.QP(Dmat_dummy, -mean_returns, Amat, bvec, meq = 1)
min_return = sum(mean_returns * result_min$solution)

# Max return
result_max = solve.QP(Dmat_dummy, mean_returns, Amat, bvec, meq = 1)
max_return = sum(mean_returns * result_max$solution)

target_return = (min_return + max_return) / 2

Dmat = 2 * cov(returns)
dvec = rep(0, n_assets)

Amat = cbind(
  mean_returns,             # target return constraint
  rep(1, n_assets),         # sum of weights = 1
  diag(n_assets)            # no shorting
)
bvec = c(target_return, 1, rep(0, n_assets))

result = solve.QP(Dmat, dvec, Amat, bvec, meq = 2)
opt_weights = result$solution
names(opt_weights) = colnames(returns)
max_sharpe_weights = opt_weights
portfolio_returns = Return.portfolio(returns, weights = opt_weights)

common_dates = index(portfolio_returns)[index(portfolio_returns) %in% index(spy_returns)]

portfolio_returns_aligned = portfolio_returns[common_dates]
spy_returns_aligned = spy_returns[common_dates]
max_sr_returns_aligned = portfolio_returns_aligned

combined_returns= merge.xts(portfolio_returns_aligned, spy_returns_aligned)
colnames(combined_returns) = c("Portfolio", "S&P 500")

chart.CumReturns(combined_returns,
                 wealth.index = T,
                 main = "Portfolio vs S&P500",
                 legend.loc = "topleft")

chart.Drawdown(combined_returns,
               legend.loc = "bottomleft",
               main = "Drawdowns")
opt_weights
SharpeRatio.annualized(combined_returns, Rf = 0)
```

```{r}
combined_returns = merge.xts(
  spy_returns_aligned,
  eq_weight_returns_aligned,
  min_var_returns_aligned,
  max_sr_returns_aligned)

colnames(combined_returns) = c(
  "S&P500",
  "Eq. Wgt. Portfolio",
  "Min. Var. Portfolio",
  "Max Sharpe Portfolio")

chart.CumReturns(combined_returns,
                 wealth.index = T,
                 main = "Portfolio vs S&P500",
                 legend.loc = "topleft")

chart.Drawdown(combined_returns,
               legend.loc = "bottomleft",
               main = "Drawdowns")

SharpeRatio.annualized(combined_returns)
```

### Of of sample performance
```{r}
from = "2024-06-01"
to = "2024-12-31"

tickers = c(
  "DELL",
  "NVDA",
  "MU", # Tech
  "GE",
  "UBER", # Industrials
  "AXP",
  "JPM", # Financials
  "AMZN",
  "HLT", # Consumer Discretionary
  "TRGP", # Energy
  "LLY",
  "BSX", # Health Care
  "NRG", # Utilities
  "BND",
  "GSG" # Fixed-Income and Commodity ETFs
)

getSymbols(tickers, from = from, to = to)
prices = do.call(merge, lapply(tickers, function(t) Ad(get(t))))
returns = na.omit(Return.calculate(prices))

tickers = c("SPY") # S&P500 as benchmark
getSymbols(tickers, from = from, to = to)
spy_prices = do.call(merge, lapply(tickers, function(t) Ad(get(t))))
spy_returns = na.omit(Return.calculate(spy_prices))
```

```{r}
eq_weight_returns = Return.portfolio(returns, weights = equal_weights)
min_var_returns = Return.portfolio(returns, weights = min_var_weights)
max_sr_returns = Return.portfolio(returns, weights = max_sharpe_weights)

combined_returns = merge.xts(
  spy_returns,
  eq_weight_returns,
  min_var_returns,
  max_sr_returns
)

colnames(combined_returns) = c(
  "S&P500",
  "Eq. Wgt. Portfolio",
  "Min. Var. Portfolio",
  "Max Sharpe Portfolio")

chart.CumReturns(combined_returns,
                 wealth.index = T,
                 main = "Portfolio vs S&P500",
                 legend.loc = "topleft")

chart.Drawdown(combined_returns,
               legend.loc = "bottomleft",
               main = "Drawdowns")

SharpeRatio.annualized(combined_returns)
```
## Sentiment Analysis
```{r}
data = read.csv("./data/Cole-sentiment-0417.csv")
data$Date <- as.Date(data$Date,format="%m/%d/%Y")

data$WeeklyHigh <- as.numeric(as.character(data$WeeklyHigh))
data$WeeklyLow <- as.numeric(as.character(data$WeeklyLow))
data$WeeklyClose <- as.numeric(as.character(data$WeeklyClose))

data$Bullish <- as.numeric(as.character(data$Bullish))
data$Neutral <- as.numeric(as.character(data$Neutral))
data$Bearish <- as.numeric(as.character(data$Bearish))


#install.packages('datetime')
library('datetime')


library(caret)
library(ggplot2)
library(dplyr)
library(tidyr)
library(mgcv)
library(crayon)
library("recipes")
library(magrittr)
library(zoo)


library(PerformanceAnalytics)
library(tidyquant)
#install.packages('forecast')
library(forecast)


####STOCK TICKER DATA MERGE###############

tickers = "META"
sentiment = data[1:4] 

colnames(sentiment) <- c("date","Bullish","Neutral","Bearish")

price_data = tq_get(
  x=tickers,get='stock.prices',
  from='1987-07-24',to='2025-04-17') 

stock <- left_join(sentiment, price_data, by = "date")
stock <- stock[complete.cases(stock), ]


returns <- ROC(stock$close, type = "discrete")
returns <- returns[-1]
#ret_demean <- returns - mean(returns)
stockminus = stock[-1,]
ret_demean <- returns
returns2 <- cbind(stockminus,ret_demean)



nrow(stock)*.75

train = returns2[1:floor(nrow(returns2)*.85),]
test = returns2[ceiling(nrow(returns2)*.85):nrow(returns2),]

plot(train$ret_demean)

last_train_value <- tail(train$close, 1)
last_test_value <- tail(test$close, 1)

############GAM MODEL#################

model2=gam(formula = ret_demean ~ Bullish + Bearish, data = returns2)
summary(model2)

pred = predict(model2,newdata=test)
rmse_train_rpart = sqrt(mean((pred[2] - test$ret_demean)^2)); rmse_train_rpart
pred <- tibble(date = test[,1], ret_demean = pred)

ggplot() +
  geom_line(data = train, aes(x = date, y = ret_demean, color = "Training"))+
  geom_line(data = test, aes(x = date, y = ret_demean, color = "Test")) +
  geom_line(data = pred, aes(x = date, y = ret_demean, color = "Forecast"))+
  scale_color_manual(name = "Data Type",values = c("Training" = "black","Test" = "red","Forecast" = "darkviolet"))+
  labs(title = "GAM Model Returns - META", x = "Date", y = "Return")+
  theme(plot.title = element_text(hjust = 0.5))

predicted_prices_simple <- last_train_value * cumprod(1 + pred[2])
predicted_price_gam <- tibble(date = test[,1], gam = predicted_prices_simple)

ggplot() +
  geom_line(data = train, aes(x = date, y = close, color = "Training"))+
  geom_line(data = test, aes(x = date, y = close, color = "Test")) +
  geom_line(data = predicted_price_gam, aes(x = date, y = gam$ret_demean, color = "Forecast"))+
  scale_color_manual(name = "Data Type",values = c("Training" = "black","Test" = "red","Forecast" = "darkviolet"))+
  labs(title = "GAM Model Price - META", x = "Date", y = "Price")+
  theme(plot.title = element_text(hjust = 0.5))

#######Gradient Boosting, gbm###########

#install.packages('gbm')
library(gbm)
boost=gbm(formula = ret_demean ~ Bullish + Bearish + Neutral, data = train,
          distribution="gaussian",
          n.trees = 100,
          interaction.depth = 3,
          shrinkage = 0.081,
          n.minobsinnode = 10)

pred_train = predict(boost, newdata = test, n.trees=100)
rmse_train_boost = sqrt(mean((pred_train - test$ret_demean)^2)); rmse_train_boost           

summary(boost)
row.names(pred_train)<- row.names(pred_train)

boostforecast<- data.frame(date = test[,1],forecast = as.numeric(pred_train))


#last_train_value <- tail(train$close, 1)
#first_predicted_value <- boostforecast[1,2]
#offset <- last_train_value - first_predicted_value
#adj_boostfore <- boostforecast[2] + offset
#adj_boostfore2 <- tibble(date = test[,1],Forecast =  adj_boostfore[,1])

ggplot() +
  geom_line(data = train, aes(x = date, y = ret_demean, color = "Training")) +
  geom_line(data = test, aes(x = date, y = ret_demean, color = "Test")) +
  geom_line(data = boostforecast, aes(x = date, y = forecast, color = "Forecast"))+
  scale_color_manual(name = "Data Type",values = c("Training" = "black","Test" = "red","Forecast" = "blue"))+
  labs(title = "Gradient Boost Model Returns - META", x = "Date", y = "Return")+
  theme(plot.title = element_text(hjust = 0.5))

predicted_prices_simple <- last_train_value * cumprod(1 + boostforecast[2])
predicted_price_gbm <- tibble(date = test[,1], gbm = predicted_prices_simple)

ggplot() +
  geom_line(data = train, aes(x = date, y = close, color = "Training"))+
  geom_line(data = test, aes(x = date, y = close, color = "Test")) +
  geom_line(data = predicted_price_gbm, aes(x = date, y = gbm$forecast, color = "Forecast"))+
  scale_color_manual(name = "Data Type",values = c("Training" = "black","Test" = "red","Forecast" = "blue"))+
  labs(title = "Gradient Boost Model Price - META", x = "Date", y = "Price")+
  theme(plot.title = element_text(hjust = 0.5))

###rpart##################


#install.packages('randomForest')
library(randomForest)
rfmodel <- randomForest(ret_demean ~ Bullish + Bearish + Neutral, data = train, ntree = 500)
summary(rfmodel)

pred_rf = predict(rfmodel, newdata = test, n.trees=100)
rmse_train_rpart = sqrt(mean((pred_rf - test$ret_demean)^2)); rmse_train_rpart 

summary(rfmodel)
varImpPlot(rfmodel)


row.names(pred_rf)<- row.names(test)

rfforecast<- data.frame(date = test[,1],forecast = as.numeric(pred_rf))


ggplot() +
  geom_line(data = train, aes(x = date, y = ret_demean, color = "Training")) +
  geom_line(data = test, aes(x = date, y = ret_demean, color = "Test")) +
  geom_line(data = rfforecast, aes(x = date, y = forecast, color = "Forecast"))+
  scale_color_manual(name = "Data Type",values = c("Training" = "black","Test" = "red","Forecast" = "darkgreen"))+
  labs(title = "Random Forest Model Returns - META", x = "Date", y = "Return")+
  theme(plot.title = element_text(hjust = 0.5))

predicted_prices_simple <- last_train_value * cumprod(1 + rfforecast[2])
predicted_price_rf <- tibble(date = test[,1], gbm = predicted_prices_simple)

ggplot() +
  geom_line(data = train, aes(x = date, y = close, color = "Training"))+
  geom_line(data = test, aes(x = date, y = close, color = "Test")) +
  geom_line(data = predicted_price_rf, aes(x = date, y = gbm$forecast, color = "Forecast"))+
  scale_color_manual(name = "Data Type",values = c("Training" = "black","Test" = "red","Forecast" = "darkgreen"))+
  labs(title = "Random Forest Model Price - META", x = "Date", y = "Price")+
  theme(plot.title = element_text(hjust = 0.5))


################AUTO ARIMA###################



train2 = stock[1:floor(nrow(stock)*.85),]
test2 = stock[ceiling(nrow(stock)*.85):nrow(stock),]


price_xts <- xts(train2$close, order.by = train2$date)
model3 = auto.arima(stock$close, lambda = 'auto', approximation = F, allowdrift = TRUE)
summary(model3)





forecast(model3,h=61)

fs <- forecast(model3,h=nrow(test))

forecast_values <- fs$mean
#row.names(forecast_values)<- row.names(test)

forecast<- data.frame(date = test[,1],forecast = as.numeric(forecast_values))

last_train_value <- tail(train2$close, 1)
first_predicted_value <- forecast[1,2]
offset <- last_train_value - first_predicted_value
AA_forecast <- forecast[2] + offset
AA_forecast2 <- tibble(date = test[,1],Forecast =  AA_forecast[,1])



ggplot() +
  geom_line(data = train2, aes(x = date, y = close, color = "Training")) +
  geom_line(data = test2, aes(x = date, y = close, color = "Test")) +
  geom_line(data = AA_forecast2, aes(x = date, y = Forecast, color = "Forecast"))+
  scale_color_manual(name = "Data Type",values = c("Training" = "black","Test" = "red","Forecast" = "blue"))+
  labs(title = "ARIMA Model - META", x = "Date", y = "Price")+
  theme(plot.title = element_text(hjust = 0.5))
```

